A community-maintained standard library of population genetic models

### 2020-05-15 popSimManu.tex

We validated our implementation of the \texttt{SLiM} engine by comparing
estimates of several population genetic summary statistics for neutral simulations generated by both \texttt{SLiM}
and \texttt{msprime}.
Examples of this validation for the AncientEurasia\_9K19 model
\citep{kamm2019efficiently}
are shown in Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}.
For this model,
down-scaling factors of up to $50$ produce patterns of both diversity and linkage disequilibrium that are
very similar to those observed under the coalescent (i.e., \texttt{msprime}), when both are simulated using the HapMapII\_GRCh37 recombination map
(Figure \ref{fig:slim_val_map}),
although the distribution of Tajima's $D$ differs slightly when scaling down by a factor of $50$.
%likely due to the subtle differences produced by down-scaling.
Interestingly, the effects of down-scaling are more prominent and extend to other sequence-based statistics
when data sets are simulated under a uniform recombination rate (Figure \ref{fig:slim_val_nomap}).
Importantly, both comparisons validate the equivalence of \texttt{SLiM} and \texttt{msprime} when no down-scaling is applied.
The results are also optimistic about the rescaling strategy to reduce computational burden,
but the possible effects are not well-understood,
so results relying on rescaled simulations should be carefully validated.

@@

\pagebreak
\beginsupplement
\section*{Supplemental Figures}
\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/HomSap_AE9K19_chr22_map_AI.pdf}
\caption{\textbf{Validating the \texttt{SLiM} engine backend under a genetic map}.
Here we validate our integration of the \texttt{SLiM} \citep{haller2019tree,haller2019slim} engine backend.
We show quantile-quantile plots between \texttt{SLiM} and \texttt{msprime} engines for three
population genetic summary statistics: \textit{r}\textsuperscript{2}, Tajima's $\pi$, and Tajima's D. Additionally, we
show runtimes for generating each simulation replicate. Data were generated by simulating 100 replicates
of human chromosome 22 under the AncientEurasia\_9K19 model \citep{kamm2019efficiently} and using the
HapMapII\_GRCh37 genetic map \citep{international2007second}. From top to bottom we show results
using three scaling factors for the population sizes: Q=1, Q=10, and Q=50.
}
\label{fig:slim_val_map}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/HomSap_AE9K19_chr22_nomap_AI.pdf}
\caption{\textbf{Validating the \texttt{SLiM} engine backend under uniform recombination}.
Here we validate our integration of the \texttt{SLiM} \citep{haller2019tree,haller2019slim} engine backend.
We show quantile-quantile plots between \texttt{SLiM} and \texttt{msprime} engines for three
population genetic summary statistics: \textit{r}\textsuperscript{2}, Tajima's $\pi$, and Tajima's D. Additionally, we
show runtimes for generating each simulation replicate. Data were generated by simulating 100 replicates
of human chromosome 22 under the AncientEurasia\_9K19 model \citep{kamm2019efficiently} and using a uniform
rate of recombination across the chromosome. From top to bottom we show results
using three scaling factors for the population sizes: Q=1, Q=10, and Q=50.
}
\label{fig:slim_val_nomap}
\end{center}
\end{figure}

>>>>

We validated our implementation of the \texttt{SLiM} engine by comparing
estimates of several population genetic summary statistics for neutral simulations generated by both \texttt{SLiM}
and \texttt{msprime}.
Examples of this validation for the AncientEurasia\_9K19 model
\citep{kamm2019efficiently}
are shown in Figures \ref{fig:slim_val_map} and \ref{fig:slim_val_nomap}.
For this model,
down-scaling factors of up to $10$ produce patterns of both diversity and linkage disequilibrium that are
indistinguishable from those observed under the coalescent (i.e., \texttt{msprime}).
Scaling down by a factor of $50$ does appear to modify the distribution of these sequence statistics.
Interestingly, the apparent difference between distributions is somewhat larger when simulating using 
a uniform recombination rate (Figure \ref{fig:slim_val_nomap}), likely due to the lower variation in
the values of these statistics.
Importantly, both comparisons validate the equivalence of \texttt{SLiM} and \texttt{msprime} when no down-scaling is applied.
The results are also optimistic about the rescaling strategy to reduce computational burden,
but the possible effects are not well-understood,
so results relying on rescaled simulations should be carefully validated.

@@

\pagebreak
\beginsupplement
\section*{Supplemental Figures}
\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/HomSap_AE9K19_chr22_map_AI.pdf}
\caption{\textbf{Validating the \texttt{SLiM} engine backend under a genetic map}.
Here we validate our integration of the \texttt{SLiM} \citep{haller2019tree,haller2019slim} engine backend.
We show quantile-quantile plots between \texttt{SLiM} and \texttt{msprime} engines for three
population genetic summary statistics: \textit{r}\textsuperscript{2}, Tajima's $\pi$, and Tajima's D. Additionally, we
show runtimes for generating each simulation replicate. Data were generated by simulating 100 replicates
of human chromosome 22 under the AncientEurasia\_9K19 model \citep{kamm2019efficiently} using the
HapMapII\_GRCh37 genetic map \citep{international2007second}. 12 samples were drawn from each population 
(excluding basal Eurasians). From top to bottom we show results
using three scaling factors for the population sizes: Q=1, Q=10, and Q=50.
Kolmogorov-Smirnov 2-sample test statistics (\textit{D}) and p-values are shown, testing the null hypothesis that
the quantiles were drawn from the same continuous distribution.
}
\label{fig:slim_val_map}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/HomSap_AE9K19_chr22_nomap_AI.pdf}
\caption{\textbf{Validating the \texttt{SLiM} engine backend under uniform recombination}.
Here we validate our integration of the \texttt{SLiM} \citep{haller2019tree,haller2019slim} engine backend.
We show quantile-quantile plots between \texttt{SLiM} and \texttt{msprime} engines for three
population genetic summary statistics: \textit{r}\textsuperscript{2}, Tajima's $\pi$, and Tajima's D. Additionally, we
show runtimes for generating each simulation replicate. Data were generated by simulating 100 replicates
of human chromosome 22 under the AncientEurasia\_9K19 model \citep{kamm2019efficiently} using a 
uniform rate of recombination across the chromosome. 12 samples were drawn from each population 
(excluding basal Eurasians). From top to bottom we show results
using three scaling factors for the population sizes: Q=1, Q=10, and Q=50.
Kolmogorov-Smirnov 2-sample test statistics (\textit{D}) and p-values are shown, testing the null hypothesis that
the quantiles were drawn from the same continuous distribution.
}
\label{fig:slim_val_nomap}
\end{center}
\end{figure}
